\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}

\title{CS 486 Assignment A (Part A)}
\author{Longling Name}
\date{\today}

\begin{document}
\maketitle


\section{Chosen problem}
We study \textbf{human--LLM collaboration for generating, repairing, and certifying feasible schedules} for a classical \emph{job-shop scheduling} style specification: each job is a fixed sequence of operations; each operation requires a machine for a known duration; each machine handles at most one operation at a time; the objective is to minimize makespan (or, in a relaxed variant, to obtain any feasible schedule with a certified feasibility check).

This problem is a concrete instance of a broader theme: \emph{NP-hard combinatorial optimization with cheap \emph{syntax} (constraint) checking but expensive \emph{semantic} (globally optimal) reasoning}.

\section{Coding platform and experiment setup}

All experiments are implemented in the public repository
\url{https://github.com/genglongling/M-APPLE-OS}.
Table~\ref{tab:main_benchmark} and Table~\ref{tab:ablation}  lays out the benchmark-by-method evaluation matrix and ablation studies.


\begin{table*}[t]
  \centering
  \small
  \setlength{\tabcolsep}{4pt}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{l|l|ccc|ccc}
  \toprule
  \textbf{Benchmark} & \textbf{Method} 
  & \multicolumn{3}{c|}{\textbf{Static Planning}} 
  & \multicolumn{3}{c}{\textbf{Dynamic / Disruption}} \\
  \cline{3-8}
  & 
  & Feasibility $\uparrow$ 
  & Violations $\downarrow$ 
  & Makespan Gap $\downarrow$
  & Replan Success $\uparrow$ 
  & Repair Cost $\downarrow$ 
  & State Errors $\downarrow$ \\
  \midrule
  
  \multirow{6}{*}{DMU} 
  & Single-Agent &  &  &  &  &  &  \\
  & Single-Agent +CoT/ReAct &  &  &  &  &  &  \\
  & Single-Agent+ Self-refine &  &  &  &  &  &  \\
  & MAS (LangGraph/AutoGen/GPTSwarm/AFlow) &  &  &  &  &  &  \\
  % HTN-style
  % reviews
  & OR/DRL &  &  &  &  &  &  \\
  & \textbf{ALAS} &  &  &  &  &  &  \\
  
  \midrule
  
  \multirow{6}{*}{TA} 
  & Single-Agent &  &  &  &  &  &  \\
  & Single-Agent +CoT/ReAct &  &  &  &  &  &  \\
  & Single-Agent+ Self-refine &  &  &  &  &  &  \\
  & MAS (LangGraph/AutoGen/GPTSwarm/AFlow) &  &  &  &  &  &  \\
  % HTN-style
  % reviews
  & OR/DRL &  &  &  &  &  &  \\
  & \textbf{ALAS} &  &  &  &  &  &  \\
  
  \midrule
  
  \multirow{6}{*}{SWZ} 
  & Single-Agent &  &  &  &  &  &  \\
  & Single-Agent +CoT/ReAct &  &  &  &  &  &  \\
  & Single-Agent+ Self-refine &  &  &  &  &  &  \\
  & MAS (LangGraph/AutoGen/GPTSwarm/AFlow) &  &  &  &  &  &  \\
  % HTN-style
  % reviews
  & OR/DRL &  &  &  &  &  &  \\
  & \textbf{ALAS} &  &  &  &  &  &  \\
  
  \midrule
  
  \multirow{6}{*}{YN} 
  & Single-Agent &  &  &  &  &  &  \\
  & Single-Agent +CoT/ReAct &  &  &  &  &  &  \\
  & Single-Agent+ Self-refine &  &  &  &  &  &  \\
  & MAS (LangGraph/AutoGen/GPTSwarm/AFlow) &  &  &  &  &  &  \\
  % HTN-style
  % reviews
  & OR/DRL &  &  &  &  &  &  \\
  & \textbf{ALAS} &  &  &  &  &  &  \\
  
  \midrule
  
  \multirow{6}{*}{ABZ} 
  & Single-Agent &  &  &  &  &  &  \\
  & Single-Agent +CoT/ReAct &  &  &  &  &  &  \\
  & Single-Agent+ Self-refine &  &  &  &  &  &  \\
  & MAS (LangGraph/AutoGen/GPTSwarm/AFlow) &  &  &  &  &  &  \\
  % HTN-style
  % reviews
  & OR/DRL &  &  &  &  &  &  \\
  & \textbf{ALAS} &  &  &  &  &  &  \\
  
  \bottomrule
  \end{tabular}
  }
  \caption{
  Main results on industrial job-shop scheduling benchmarks (DMU, TA, SWZ, YN, ABZ).
  Static planning evaluates feasibility and optimization quality, while dynamic settings assess disruption recovery,
  state consistency, and repair efficiency. ALAS is expected to achieve the strongest performance under dynamic disruptions
  while remaining competitive in static optimization.
  }
  \label{tab:main_benchmark}
  \end{table*}

  \begin{table}[t]
    \centering
    \small
    \setlength{\tabcolsep}{4pt}
    \resizebox{\columnwidth}{!}{
    \begin{tabular}{lcccccc}
    \toprule
    \textbf{Variant} 
    & \textbf{Horizon} 
    & \textbf{State} 
    & \textbf{Constraint} 
    & \textbf{Replan} 
    & \textbf{Repair Cost} 
    & \textbf{Robustness} \\
    \midrule
    
    ALAS (full) 
    & \textbf{High} 
    & \textbf{High} 
    & \textbf{High} 
    & \textbf{High} 
    & \textbf{Low} 
    & \textbf{High} \\
    
    w/o Memory 
    & Medium 
    & Low 
    & Medium 
    & Low 
    & High 
    & Low \\
    
    w/o Validator 
    & Medium 
    & Medium 
    & Low 
    & Medium 
    & Medium 
    & Medium \\
    
    w/o LRCP 
    & Medium 
    & Medium 
    & Medium 
    & Low 
    & High 
    & Low \\
    
    % Single-Agent 
    % & Low 
    % & Low 
    % & Low 
    % & Low 
    % & High 
    % & Low \\
    
    \bottomrule
    \end{tabular}
    }
    \caption{
    Ablation study of ALAS components across key planning capabilities. 
    Removing memory degrades state consistency, removing validation harms constraint satisfaction, 
    and removing LRCP significantly reduces replanning effectiveness and increases repair cost.
    }
    \label{tab:ablation}
    \end{table}

\section{Related work}

\paragraph{Job-shop scheduling and complexity.}
The job-shop problem to minimize makespan is a standard strongly NP-hard model in machine scheduling~\cite{GJ79,LRKB77}.
Exact methods (branch-and-bound, constraint programming, mixed-integer programming) and dedicated JSSP algorithms have been studied for decades; representative exact/heuristic lines include the shifting bottleneck heuristic~\cite{Adams1988}, branch-and-bound frameworks for the job-shop problem~\cite{Carlier1989}, and high-performance tabu search~\cite{Nowicki1996}.
Textbooks and monographs synthesize complexity, dispatching rules, and metaheuristics for production scheduling~\cite{CPHandbook}.

\paragraph{Large language models as collaborators (not certified solvers).}
Foundation models are widely used for step-by-step reasoning prompts (chain-of-thought)~\cite{Wei2022} and for interleaving reasoning with tool use (ReAct-style agents)~\cite{Yao2023}.
Multi-agent \emph{conversation} frameworks such as AutoGen coordinate roles and tool calls among LLMs and humans~\cite{Wu2024}.
In operations research, there is peer-reviewed work on using language models to help translate natural-language problem descriptions into formal optimization formulations, with human validation in the loop~\cite{Ramamonjison2022}.
These lines support using LLMs for modeling, explanation, and proposal generation, but they do not replace the need for independent verification of feasibility and optimality on concrete instances.

\paragraph{Gap addressed in this project.}
Classical scheduling literature assumes solutions are produced by specified algorithms; recent LLM agent work emphasizes flexible dialogue but often elides \emph{deterministic} checking of constraints.
Our setup follows the OR ``augmentation'' perspective---LLM proposals plus human oversight~\cite{Ramamonjison2022}---while tightening the loop with explicit verifiers, logged failures, and benchmark artifacts (Section~\ref{sec:problem-spec}).

\section{Problem specification}
\label{sec:problem-spec}

\paragraph{Survey methodology.} We used an LLM as a \emph{guidance and summarization} assistant, not as an oracle:
\begin{enumerate}[noitemsep]
  \item Seed queries with precise keywords (\texttt{job shop NP-hard}, \texttt{disjunctive graph}, \texttt{branch and bound JSSP}, \texttt{large neighborhood search scheduling}, \texttt{LLM operations research scheduling}).
  \item Require the LLM to return \emph{search strings}, venues, and author names; we then opened primary sources (journal PDFs, textbooks, reputable surveys).
  \item For each ``theorem-level'' claim extracted by the LLM, we checked: statement, hypotheses, and whether the source is a proof or an empirical report.
\end{enumerate}

\textbf{Known results (high level).}
\begin{itemize}[noitemsep]
  \item \textbf{Computational complexity.} The job-shop problem to minimize makespan is strongly NP-hard; exact methods (branch-and-bound, constraint programming, MIP) scale poorly as jobs and machines grow.
  \item \textbf{Classical algorithmic cores.} Industrial practice relies on dispatching rules, metaheuristics (tabu, genetic algorithms, simulated annealing), and large neighborhood search; these often lack \emph{universal} worst-case guarantees matching the best-known approximation bounds for restricted variants.
  \item \textbf{Learning and LLMs in scheduling.} Recent work explores LLMs for modeling, code synthesis, and heuristic design; however, \emph{plan feasibility} and \emph{plan constraint satisfaction} remain only checkable by traditional solvers, while \emph{plan optimality} and \emph{plan scalability} are generally unavailable for large instances within strict time budgets.
\end{itemize}

\subsection{Our Method: ALAS Workflow}
We decompose the collaboration target into four layers:
\begin{enumerate}[label=\textbf{L\arabic*.}, leftmargin=*]
  \item \textbf{Problem formalization layer.} Parse instance data (jobs, machines, durations, routes) into a structured representation suitable for both human inspection and automated checking.
  \item \textbf{Candidate generation layer.} Propose a schedule (start/end times per operation) using heuristics, LLM-authored code, or hybrid search; this layer is creative and error-prone.
  \item \textbf{Verification layer.} Check precedence, machine capacity, and completeness (all operations scheduled exactly once) with deterministic procedures independent of the LLM.
  \item \textbf{Improvement layer.} If verification fails, localize violations and repair (human-guided or algorithmic); if verification succeeds, optionally run optimization or bound computation as a separate, clearly scoped step.
\end{enumerate}

\subsection{Hard core}
\begin{itemize}[noitemsep]
  \item \textbf{Tractable (relative to human + modest compute).} Feasibility checking for a \emph{given} schedule; small-scale exact optimization; bounded local repair moves; logging and diffing of intermediate artifacts; reproducible scripts.
  \item \textbf{Hard core.} Global optimality (or tight approximation guarantees) for large unstructured instances within strict time limits; reliably \emph{generating} feasible high-quality schedules without search; any claim of ``this is optimal'' without an explicit certificate from a trusted solver or a proved special-case theorem.
\end{itemize}

\subsection{Planned angle of attack}
Our angle is \textbf{separation of concerns}: treat the LLM as a \emph{proposal engine} and possibly as an \emph{explainer} of repair strategies, but treat \emph{truth} about feasibility as the output of Layer~L3 only. Concretely:
\begin{itemize}[noitemsep]
  \item Standardize on a machine-readable schedule format and a single verifier implementation (the ``ground truth'' for feasibility).
  \item Iterate \textbf{generate $\rightarrow$ verify $\rightarrow$ diagnose $\rightarrow$ repair}, with explicit logging of failed attempts (counteracting memory loss across sessions).
  \item Reserve claims of optimality for instances where an exact solver or a clearly stated relaxation bound is available; otherwise report \emph{makespan} with the status \emph{feasible / infeasible / unknown optimality}.
\end{itemize}

\section{Protocol design}

\subsection{Using CS486 ``knobs'' explicitly}

\paragraph{Knob 1: Hypothesis discipline (vs.\ exaggeration/confabulation).}
\textbf{Use:} Every substantive claim is tagged in the shared log as one of:
\textsc{proved} (with citation and exact theorem statement),
\textsc{checked-by-code} (script name + input hash + output),
\textsc{empirical} (instance class + experimental protocol),
\textsc{conjecture} (explicitly labeled).
\textbf{Prompt pattern:} ``State the \emph{minimal} hypotheses needed. If you cannot prove it, move the statement to \textsc{conjecture} and list what would falsify it.''

\paragraph{Knob 2: Adversarial review (vs.\ sycophancy).}
\textbf{Use:} The Critic's success metric is \emph{finding} issues, not agreement.
\textbf{Prompt pattern:} ``Your job is to break this schedule or argument. If you agree, you must still spend 200 words on the strongest counter-case. Then argue the opposite position for one paragraph.''
\textbf{Test:} If the model flips from ``sound'' to ``fatally flawed'' with a mild rewording of the user hypothesis, treat prior endorsement as unreliable.

\paragraph{Knob 3: External memory (vs.\ no regret / memory loss).}
\textbf{Use:} A human-maintained \texttt{collab\_log.md} (or equivalent) with sections:
\texttt{FAILED\_APPROACHES}, \texttt{RETRACTED\_CLAIMS}, \texttt{VERIFIED\_FACTS}, \texttt{OPEN\_QUESTIONS}.
At the start of each session, paste the log into context \emph{before} asking for new ideas.
\textbf{Rule:} If an approach appears twice without addressing why it failed before, stop and require a diff against the prior failure.

\paragraph{Knob 4: Paradigm pivots (vs.\ paradigm blindness).}
\textbf{Use:} When progress stalls, the human explicitly forces a framework shift (e.g., ``encode as CP-SAT,'' ``use disjunctive graph visualization,'' ``reduce to maximum clique on interval conflicts,'' ``switch from narrative proof to counterexample search'').
\textbf{Prompt pattern:} ``Do not refine the previous approach. Pick a different mathematical packaging and restate the problem in that language only.''


\subsection{Roles and information flow}
We use a small multi-role pattern (these may be separate prompts, separate agents, or the same model with a \emph{forced role switch}):
\begin{itemize}[noitemsep]
  \item \textbf{Architect:} states assumptions, decomposes the instance, proposes plan of attack.
  \item \textbf{Builder:} emits structured schedule artifacts and code.
  \item \textbf{Critic:} adversarial review; must output \emph{falsifiable} objections (e.g., a specific operation overlap).
  \item \textbf{Verifier (non-LLM):} deterministic checker.
\end{itemize}


\subsection{The four LLM failure modes: concrete countermeasures}

\paragraph{1. Exaggeration and confabulation.}
\begin{itemize}[noitemsep]
  \item Require scope tags on claims (\textsc{proved} / \textsc{checked-by-code} / \textsc{empirical} / \textsc{conjecture}).
  \item For anything that ``sounds like a theorem,'' ask: \emph{Which sentence uses which hypothesis?} and demand a line-by-line map.
  \item Never accept ``this algorithm is optimal'' without a certificate source named in the log.
\end{itemize}

\paragraph{2. Sycophancy.}
\begin{itemize}[noitemsep]
  \item Separate Builder and Critic prompts; instruct Critic to maximize defect detection.
  \item Periodic \emph{red-team} prompts: argue against the user's favorite approach.
  \item Cross-check with non-LLM tools (verifier, solver) whenever agreement appears ``too easy.''
\end{itemize}

\paragraph{3. No persistent memory, no regret.}
\begin{itemize}[noitemsep]
  \item Running \texttt{FAILED\_APPROACHES} and \texttt{RETRACTED\_CLAIMS} lists, re-injected each session.
  \item Versioned artifacts (schedules, logs) with hashes/timestamps; require the LLM to reference the latest artifact filename when proposing edits.
\end{itemize}

\paragraph{4. No intuition, no paradigm bridging.}
\begin{itemize}[noitemsep]
  \item Human-owned ``pivot checklist'' (CP/MIP, graph formulations, probabilistic intuition, local search neighborhoods).
  \item When stuck, mandate a \emph{new} representation before more local edits.
\end{itemize}

\subsection{Verification strategy (end-to-end)}
\textbf{Primary verifier (trusted):} a deterministic checker that validates:
(i) each job's operations respect precedence and durations;
(ii) each machine has no overlapping operations;
(iii) every operation appears exactly once.
\textbf{Secondary checks (optional):} compare makespan to known upper bounds or best-known values for benchmark instances; run a small exact solver on a reduced sub-instance as a sanity check.
\textbf{Reporting rule:} publish \emph{feasibility verdict} from the primary verifier; publish \emph{optimality} only when a named exact method certifies it or a proved bound applies.

\subsection{Extension to JSSP-dynamic}
Because scheduling outputs can affect real resource allocation, the protocol also logs \emph{assumptions about uncertainty} (durations known vs.\ stochastic) and avoids presenting heuristic outputs as certifiably safe without domain review.

\begin{thebibliography}{99}

\bibitem{GJ79}
M.~R. Garey and D.~S. Johnson.
\newblock \emph{Computers and Intractability: A Guide to the Theory of NP-Completeness}.
\newblock W.~H. Freeman, 1979.

\bibitem{LRKB77}
J.~K. Lenstra, A.~H.~G. Rinnooy Kan, and P. Brucker.
\newblock Complexity of machine scheduling problems.
\newblock \emph{Annals of Discrete Mathematics}, 1:343--362, 1977.

\bibitem{CPHandbook}
M.~L. Pinedo.
\newblock \emph{Scheduling: Theory, Algorithms, and Systems}.
\newblock Springer Science+Business Media, 2008.
\newblock \href{https://doi.org/10.1007/978-0-387-78935-4}{doi:10.1007/978-0-387-78935-4}.

\bibitem{Adams1988}
J.~Adams, E.~Balas, and D.~Zawack.
\newblock The shifting bottleneck procedure for job shop scheduling.
\newblock \emph{Management Science}, 34(3):391--401, 1988.

\bibitem{Carlier1989}
J.~Carlier and E.~Pinson.
\newblock An algorithm for solving the job-shop problem.
\newblock \emph{Management Science}, 35(11):1648--1655, 1989.

\bibitem{Nowicki1996}
E.~Nowicki and C.~Smutnicki.
\newblock A fast tabu search algorithm for the job shop problem.
\newblock \emph{Management Science}, 42(6):797--813, 1996.

\bibitem{Wei2022}
J.~Wei, X.~Wang, D.~Schuurmans, M.~Bosma, B.~Ichter, F.~Xia, E.~Chi, Q.~Le, and D.~Zhou.
\newblock Chain-of-thought prompting elicits reasoning in large language models.
\newblock In \emph{Advances in Neural Information Processing Systems (NeurIPS)}, volume~35, 2022.

\bibitem{Yao2023}
S.~Yao, J.~Zhao, D.~Yu, N.~Du, I.~Shafran, K.~Narasimhan, and Y.~Cao.
\newblock {ReAct}: Synergizing reasoning and acting in language models.
\newblock In \emph{International Conference on Learning Representations (ICLR)}, 2023.

\bibitem{Wu2024}
Q.~Wu, G.~Bansal, J.~Zhang, Y.~Wu, B.~Li, E.~Zhu, L.~Jiang, X.~Zhang, S.~Zhang, J.~Liu, et al.
\newblock {AutoGen}: Enabling next-gen {LLM} applications via multi-agent conversations.
\newblock In \emph{First Conference on Language Modeling (COLM)}, 2024.
\newblock \url{https://openreview.net/forum?id=BAakY1hNKS}.

\bibitem{Ramamonjison2022}
R.~Ramamonjison, H.~Li, T.~Yu, S.~He, V.~Rengan, A.~Banitalebi-dehkordi, Z.~Zhou, and Y.~Zhang.
\newblock Augmenting operations research with auto-formulation of optimization models from problem descriptions.
\newblock In Y.~Li and A.~Lazaridou (eds.), \emph{Proceedings of the 2022 Conference on Empirical Methods in Natural Language Processing: Industry Track}, pages 29--62, Abu Dhabi, UAE, December 2022. Association for Computational Linguistics.
\newblock \href{https://doi.org/10.18653/v1/2022.emnlp-industry.4}{doi:10.18653/v1/2022.emnlp-industry.4}.
\newblock \url{https://aclanthology.org/2022.emnlp-industry.4/}.

\end{thebibliography}

\end{document}
